Comparative analysis of PMV Models accuracy implemented in the ISO 7730:2005 and ASHRAE 55:2023

### 2025-01-16 Manuscript/src/sections/methodology.tex
review: calculated Icl,r ISO, reviewed the paper until the start methodology.tex, changed V to Vr, Icl to Icl,r

\subsection{Data Preparation and Cleaning}\label{subsec:data-processing-and-cleaning}
Not all researchers measured and logged all six parameters needed to calculate the \ac{pmv} and \ac{pmv-ce}.
We, therefore, decided to remove all entries that did not have any of the following variables: \ac{tdb}, \ac{tr}, \ac{rh}, \ac{v}, \ac{clo}, \ac{met}, and \ac{tsv}.
This filtered out approximately \var{entries_db_valid} of the data.
In our analysis, we used the value \ac{v} which is adjusted as a function of \ac{met} and not the air speed measured by the researchers.
This is because both \ac{pmv} formulations use the \ac{v} adjusted as a function of \ac{met} to calculate the \ac{pmv} value.

Both the \gls{55} and \gls{7730} specify a set of applicability limits.
We hence filtered out those data points that did not meet the inclusion criteria of both standards.
The rationale is that the models' accuracy should only be tested within their applicability ranges.
The limits we used are as follows:
\num{10}~$\leq$~\ac{tdb}~$\leq$~\qty{30}{\celsius},
\num{10}~$\leq$~\ac{tr}~$\leq$~\qty{40}{\celsius},
\num{0}~$\leq$~\ac{v}~$\leq$~\qty{1}{\m\per\s},
\num{0}~$\leq$~\ac{clo}~$\leq$~\qty{1.5}{clo},
\num{0}~$\leq$~water vapor partial pressure~$\leq$~\qty{2000}{\pascal},
and \num{1}~$\leq$~\ac{met}~$\leq$~\qty{4}{met}.
These criteria removed an additional \var{entries_db_filtered_by_limit_inputs} datapoints.
Consequently, the final dataset used for the analysis contains \var{entries_db_used} entries.
The adjusted total clothing insulation and relative airspeed, as required by both standards, were used to calculate the \ac{pmv} and \ac{pmv-ce} values.

@@

In addition, it should also be noted that while all the data included in the \ac{db2} are published in peer-reviewed papers, not all entries were collected using the same methodology.
For example, while both thermal comfort standards recommend measuring \ac{v} at three different heights, this was not always done.
This could introduce a bias in the measurements and, consequently, in the results of the models.
Moreover, since the two \ac{pmv} formulations only differ when \ac{v}~$\geq$~\qty{0.1}{\m\per\s}, we decided to report three sets of results:
\begin{enumerate}[ {}1{)} ]
    \item all the data points with \ac{tdb}, \ac{tr}, \ac{rh}, \ac{v}, \ac{clo}, \ac{met}, and \ac{tsv};
    \item the data points in 1) with \ac{v}~$\geq$~\qty{0.2}{\m\per\s};
    \item the data points in 1) with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured air speed at three different heights.
    These results are reported in \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}.
\end{enumerate}
Comparing the results from these three datasets will allow us to assess the impact of the data quality on the results of the models.
By only including those studies that measured air speed at three different heights, we can reduce the bias introduced by the lack of adherence to the measurement protocol recommended by both standards.
We did not use \qty{0.1}{\m\per\s} as a threshold because the original difference between the two models was \qty{0.2}{\m\per\s} and it was changed to \qty{0.1}{\m\per\s} only in the \gls{55}.
The effect of air movement is perceptible only when the airspeed is sufficiently high to disrupt the body's thermal plume~\cite{zukowska_impact_2012}.
The threshold in the ASHRAE standard was reduced to \qty{0.1}{\m\per\s} to simplify the usability and achieve one threshold for all the air speeds.
In \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}, we also present the results generated using those entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured \ac{v} at three different heights.
We have not included these results here, as they are comparable to those presented in this section and lead to the same conclusions.

@@

To remove this rounding error, we subtracted the \ac{tsv} value from the \ac{pmv} and \ac{pmv-ce} values.
These differences, also known as bias, quantify the success of the model in predicting \ac{tsv}.
However, on their own, are a low-precision estimate of the overall accuracy of the model~\cite{Humphreys2002}.
These values were then binned using several independent variables (e.g., \ac{tdb}, \ac{v}).
If the \ac{pmv} or \ac{pmv-ce} formulations are bias-free, the distribution of any batch derived from these differences would have a mean value that is zero.
The standard deviation would reflect the combined effect of the people's individual differences, any errors in the model formulation or, in the data collection method (accuracy or precision of the instrumentation used)~\cite{Humphreys2002}.
According to \mycite{Humphreys2002} the model can be considered accurate if the above-mentioned differences are between \num{-0.25} and \num{0.25}.
However, we believe that this range is too narrow since \ac{tsv} is commonly measured using an ordinal scale of integers.
Consequently, we expanded the range to \num{-.5} and \num{0.5}.
This is needed to ensure that each \ac{pmv} value can be matched to a \ac{tsv}.
We are aware that an error of \num{.5} is approximately \qty{7}{\percent} of the total range and ideally, we would want to have the \ac{pmv} to have higher precision, however, this is a necessary limitation introduced by the low resolution of the \ac{tsv} scale.
When the aim is to predict the average thermal sensation of a large group of occupants, binning data by the independent variables is a better representation.
Grouping the participants' responses by their \ac{tsv} mixes people who reported the same \ac{tsv} vote but could have been in different locations and moments of the year and day.

>>>>

\subsection{Data Preparation and Cleaning}\label{subsec:data-processing-and-cleaning}
Not all researchers measured and logged all six parameters needed to calculate the \ac{pmv} and \ac{pmv-ce}.
We, therefore, decided to remove all entries that did not have any of the following variables: \ac{tdb}, \ac{tr}, \ac{rh}, \ac{v}, \ac{clo}, \ac{met}, and \ac{tsv}.
This filtered out approximately \var{entries_db_valid} of the data.
In our analysis, we used the value \ac{vr} which is adjusted as a function of \ac{met} and not the air speed measured by the researchers.
This is because both \ac{pmv} formulations use the \ac{vr} adjusted as a function of \ac{met} to calculate the \ac{pmv} value.

Both the \gls{55} and \gls{7730} specify a set of applicability limits.
We hence filtered out those data points that did not meet the inclusion criteria of both standards.
The rationale is that the models' accuracy should only be tested within their applicability ranges.
The limits we used are as follows:
\num{10}~$\leq$~\ac{tdb}~$\leq$~\qty{30}{\celsius},
\num{10}~$\leq$~\ac{tr}~$\leq$~\qty{40}{\celsius},
\num{0}~$\leq$~\ac{vr}~$\leq$~\qty{1}{\m\per\s},
\num{0}~$\leq$~\ac{clo}~$\leq$~\qty{1.5}{clo},
\num{0}~$\leq$~water vapor partial pressure~$\leq$~\qty{2000}{\pascal},
and \num{1}~$\leq$~\ac{met}~$\leq$~\qty{4}{met}.
These criteria removed an additional \var{entries_db_filtered_by_limit_inputs} datapoints.
Consequently, the final dataset used for the analysis contains \var{entries_db_used} entries.
The adjusted total clothing insulation and relative airspeed, as required by both standards, were used to calculate the \ac{pmv} and \ac{pmv-ce} values.

@@

In addition, it should also be noted that while all the data included in the \ac{db2} are published in peer-reviewed papers, not all entries were collected using the same methodology.
For example, while both thermal comfort standards recommend measuring \ac{v} at three different heights, this was not always done.
This could introduce a bias in the measurements and, consequently, in the results of the models.
Moreover, since the two \ac{pmv} formulations only differ when \ac{vr}~$\geq$~\qty{0.1}{\m\per\s}, we decided to report three sets of results:
\begin{enumerate}[ {}1{)} ]
    \item all the data points with \ac{tdb}, \ac{tr}, \ac{rh}, \ac{v}, \ac{clo}, \ac{met}, and \ac{tsv};
    \item the data points in 1) with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s};
    \item the data points in 1) with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured air speed at three different heights.
    These results are reported in \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}.
\end{enumerate}
Comparing the results from these three datasets will allow us to assess the impact of the data quality on the results of the models.
By only including those studies that measured air speed at three different heights, we can reduce the bias introduced by the lack of adherence to the measurement protocol recommended by both standards.
We did not use \qty{0.1}{\m\per\s} as a threshold because the original difference between the two models was \qty{0.2}{\m\per\s} and it was changed to \qty{0.1}{\m\per\s} only in the \gls{55}.
The effect of air movement is perceptible only when the airspeed is sufficiently high to disrupt the body's thermal plume~\cite{zukowska_impact_2012}.
The threshold in the ASHRAE standard was reduced to \qty{0.1}{\m\per\s} to simplify the usability and achieve one threshold for all the air speeds.
In \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}, we also present the results generated using those entries with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured \ac{v} at three different heights.
We have not included these results here, as they are comparable to those presented in this section and lead to the same conclusions.

@@

To remove this rounding error, we subtracted the \ac{tsv} value from the \ac{pmv} and \ac{pmv-ce} values.
These differences, also known as bias, quantify the success of the model in predicting \ac{tsv}.
However, on their own, are a low-precision estimate of the overall accuracy of the model~\cite{Humphreys2002}.
These values were then binned using several independent variables (e.g., \ac{tdb}, \ac{vr}).
If the \ac{pmv} or \ac{pmv-ce} formulations are bias-free, the distribution of any batch derived from these differences would have a mean value that is zero.
The standard deviation would reflect the combined effect of the people's individual differences, any errors in the model formulation or, in the data collection method (accuracy or precision of the instrumentation used)~\cite{Humphreys2002}.
According to \mycite{Humphreys2002} the model can be considered accurate if the above-mentioned differences are between \num{-0.25} and \num{0.25}.
However, we believe that this range is too narrow since \ac{tsv} is commonly measured using an ordinal scale of integers.
Consequently, we expanded the range to \num{-.5} and \num{0.5}.
This is needed to ensure that each \ac{pmv} value can be matched to a \ac{tsv}.
We are aware that an error of \num{.5} is approximately \qty{7}{\percent} of the total range and ideally, we would want to have the \ac{pmv} to have higher precision, however, this is a necessary limitation introduced by the low resolution of the \ac{tsv} scale.
When the aim is to predict the average thermal sensation of a large group of occupants, binning data by the independent variables is a better representation.
Grouping the participants' responses by their \ac{tsv} mixes people who reported the same \ac{tsv} vote but could have been in different locations and moments of the year and day.

### 2025-01-16 Manuscript/src/sections/results.tex
review: calculated Icl,r ISO, reviewed the paper until the start methodology.tex, changed V to Vr, Icl to Icl,r

\section{Results and Discussion}\label{sec:results}
While the \ac{db2} contains \var{entries_db_all} data points, only \var{entries_db_used} entries meet the inclusion criteria listed in the Methodology section.
The distribution of the six input variables used to calculate the PMV indices is visualized in Figure~\ref{fig:dist_input_data}.
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{figures/dist_input_data}
    \caption{Distribution of the input variables used to calculate the \ac{pmv} and \ac{pmv-ce} values.
    The data are shown using boxen plots (letter-value plots).
    They depict the median as the center-line and each successive level outward contains half of the remaining data.}
    \label{fig:dist_input_data}
\end{figure*}
The input data distributions are far from uniformity (equal density) across the standards' applicability ranges.
We were not able to determine the accuracy of the \ac{pmv} and \ac{pmv-ce} formulations for inputs that are lower than the 2.5th and 97.5th percentiles, since we only had a limited number of data points beyond those values.
For example, only \qty{2.5}{\percent} of the available dataset had a \ac{tdb}~$\leq$ than \var{ta_95_perc_min}, this is a problem since both thermal comfort standards' applicability limits extend to \qty{10}{\celsius}.
This can be explained by the fact that only a few buildings in the world operate at \ac{tdb} lower than this value.
Mainly because an indoor \ac{tdb} lower than \var{ta_95_perc_min} would be deemed to be unsatisfactory by a great number of occupants~\cite{iso7730}.
This lack of data in the \ac{db2} is particularly relevant for \ac{v}.
The \ac{pmv} and \ac{pmv-ce} only differ when the value of \ac{v}~$\geq$~\qty{0.1}{\m\per\s}.
The median value in the \ac{db2} is \var{v_median} suggesting that most of the data points were collected in buildings with low air movement.
We also could not reliably test the accuracy of the \ac{pmv} formulations for values of \ac{v}~$\geq$~\var{v_95_perc_max}.

@@

\subsection{Comparison of PMV Accuracy in Predicting Thermal Sensation}\label{subsec:model-accuracy-comparison-in-predicting-thermal-sensation}
The \ac{pmv} was developed with the primary aim of predicting \ac{tsv}, consequently, we compared the predicted value from the two models with measured \ac{tsv}.
However, since the two models only differ when \ac{v}~$>$~\qty{0.1}{\m\per\s};
we also present the results for a subset of the data where \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
In this subset of data, the differences in the outputs of the two models are more pronounced.
We set the limit to \qty{0.2}{\m\per\s} and not a higher value since, as previously shown in Figure~\ref{fig:dist_input_data}, we could only reliably test the accuracy of the models up to \ac{v} of \var{v_95_perc_max}.

Being \ac{tsv} the value reported by participants (i.e., ground truth), in Figure~\ref{fig:bar_stacked_model_accuracy} we grouped participants' responses by \ac{tsv} and reported the simple accuracy of the different \ac{pmv} formulations over each sub-figure.
\begin{figure}[htb!]
    \centering
    \begin{subfigure}[b]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/bar_stacked_model_accuracy_0}
    \end{subfigure}
    \par\bigskip % force a bit of vertical whitespace
    \begin{subfigure}[b]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/bar_stacked_model_accuracy_0.2}
    \end{subfigure}
    \caption{The calculated \ac{pmv} value is grouped by the \ac{tsv} value reported by the participants.
    In a) and b) for the entire dataset, while in c and d for the dataset when \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
    In a) and c) are shown the results for the \ac{pmv} model, and in b) and d) for the \ac{pmv-ce} model.
    The number of data points in each bin of the two charts at the top (a and b) and at the bottom (c and d) is the same, and this allows us to compare the accuracy of both models.
    The overall prediction accuracy is shown above the chart, while the model accuracy of predicting participants grouped by their \ac{tsv} is shown over each bar.
    A random model would have an accuracy of \qty{14}{\percent}.}
    \label{fig:bar_stacked_model_accuracy}
\end{figure}
Both \ac{pmv} formulations underestimate the conditions at which participants are thermally dissatisfied with their environment.
The overall accuracies of the \ac{pmv} and \ac{pmv-ce} are \qty{33}{\percent} and \qty{34}{\percent}, respectively, when using the whole dataset.
The overall accuracy of both \ac{pmv} models increase to \qty{34}{\percent} and \qty{35}{\percent}, respectively when only considering entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
The \ac{pmv-ce} model outperforms the \ac{pmv} model for \ac{tsv}~=~0 and -1.
However, the \ac{pmv} model outperforms the \ac{pmv-ce} model when considering \ac{tsv}~=~1.
The number of data points in each \ac{tsv} bin is different.
\mycite{Cheung2019} grouped people based on the same heat loss or gains (same \ac{pmv} ranges), not the same \ac{tsv} as we did here, and then quantified the sensitivity of the model, i.e., how many times the \ac{pmv} model correctly predicted \ac{tsv}.
They found that the accuracy of \ac{pmv} in predicting \ac{tsv} was only \qty{34}{\percent} \mycite{Cheung2019}.

@@

In Table~\ref{tab:f1}, we report the F1 scores for both models, i) using the whole dataset, ii) only keeping entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}, iii) only keeping data points that have \ac{v} measured at three heights with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}, and iv) for those with a $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
\begin{table}[htb!]
    \centering
    \input{tables/f1}
    \caption{F1 scores for the \ac{pmv} and \ac{pmv-ce} models for different subsets of data.}
    \label{tab:f1}
\end{table}
The F1-macro score, which is free from label imbalance, depicts that the accuracy of both models across all metrics is marginally better than random guessing (i.e., \qty{14}{\percent}) for all data.
For the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} the F1-macro scores of the \ac{pmv} and \ac{pmv-ce} models only slightly improved.
We then analyzed the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and with \ac{v} measured at three heights.
Both F1-macro scores decrease and the \ac{pmv-ce} (\num{0.16}) is lower than the value for the \ac{pmv}~ (\num{0.17}).
Our results show that the assumption that the \ac{pmv-ce} better takes into account the effect of elevated air speed is false.
All the F1 scores significantly increase when analyzing the dataset with $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
This highlights that the \ac{pmv} models are only accurate in predicting the thermal sensation of people when the thermal sensation is close to neutral.

@@

\subsubsection{Model Overall Bias}\label{subsubsec:model-overall-bias}
We calculated the overall bias of the model as previously done by \mycite{Humphreys2002} by subtracting the \ac{pmv} and \ac{pmv-ce} models prediction from the self-reported \ac{tsv}.
Figure~\ref{fig:hist_discrepancies} contains the summary statistics.
Overall both the \ac{pmv} and \ac{pmv-ce} models have a median bias lower than half of a thermal sensation interval.
The results highlight that the \ac{pmv-ce} has a higher bias than the \ac{pmv} both when calculating the bias for all the data and for the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
We report the median and inter-quantile ranges since data are not normally distributed.
%  We cannot change into a letter value plot since in that case, it would be hard to highlight the number of responses that are within the range from -.5 to 5.
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{figures/hist_discrepancies}
    \caption{Overall bias of the \ac{pmv} and \ac{pmv-ce} models for the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
    The x-axis shows the discrepancy (bias) between the predicted and self-reported \ac{tsv}.
    We have binned the data into intervals of equal width.
    We colored the bars in the range from \num{-.5} to \num{.5} to highlight the number of responses that are within this range.
    The chart also shows the median and inter-quantile ranges of the data.
    }
    \label{fig:hist_discrepancies}
\end{figure*}
When including only the data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} the median bias of the \ac{pmv} model reduces to \var{bias_median_pmv_0.2} from \var{bias_median_pmv_0}, meaning that at higher speed the \ac{pmv} model improves.
The median bias for the \ac{pmv-ce} also decreases to \var{bias_median_pmv_ce_0.2} from \var{bias_median_pmv_ce_0} but it is higher than the one for the \ac{pmv}.

\subsubsection{Model Bias as Function of Five Input Variables}\label{subsubsec:model-bias-variable}
Here we present the model bias as a function of the five input variables \ac{tdb}, \ac{v}, \ac{rh}, \ac{met}, and \ac{clo}.
Results are shown in Figure~\ref{fig:bias_models}.
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{figures/bias_models}
    \caption{Bias of the \ac{pmv} and \ac{pmv-ce} models plotted as a function of the input variables.
    The x ticks labels show the bin range.
    The square bracket($]$) means that the right value is included in the range.
    The red dashed lines show the acceptable range of bias.
    Each bin has a different number of data points.
    }
    \label{fig:bias_models}
\end{figure*}
In each sub-figure, we have binned the results into discrete intervals.
The x-axis tick label shows the bin range.
All the boxen-plots have the same area to aid visualization, even though the number of points in each bin is not constant.
The results depict that the use of the \ac{pmv-ce} model does not improve the bias of the \ac{pmv} results across all input variables.
On the contrary, the bias of the \ac{pmv-ce} diverges more from 0 and is significantly worse (\num{-.33} versus \num{-0.13}) than the one for the \ac{pmv} for \qty{0.3}{\m\per\s}~$<$~\ac{v}~$\leq$~\qty{0.46}{\m\per\s}.
Both models have a median bias higher or equal to $\lvert0.5\lvert$ for \ac{tdb}~$\leq$~\qty{21}{\celsius} and \ac{clo}~$\leq$~\qty{.5}{clo}.

>>>>

\section{Results and Discussion}\label{sec:results}
While the \ac{db2} contains \var{entries_db_all} data points, only \var{entries_db_used} entries meet the inclusion criteria listed in the Methodology section.
The distribution of the six input variables used to calculate the PMV indices is visualized in Figure~\ref{fig:dist_input_data}.
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{figures/dist_input_data}
    \caption{Distribution of the input variables used to calculate the \ac{pmv} and \ac{pmv-ce} values.
    The data are shown using boxen plots (letter-value plots).
    They depict the median as the center-line and each successive level outward contains half of the remaining data.}
    \label{fig:dist_input_data}
\end{figure*}
The input data distributions are far from uniformity (equal density) across the standards' applicability ranges.
We were not able to determine the accuracy of the \ac{pmv} and \ac{pmv-ce} formulations for inputs that are lower than the 2.5th and 97.5th percentiles, since we only had a limited number of data points beyond those values.
For example, only \qty{2.5}{\percent} of the available dataset had a \ac{tdb}~$\leq$ than \var{ta_95_perc_min}, this is a problem since both thermal comfort standards' applicability limits extend to \qty{10}{\celsius}.
This can be explained by the fact that only a few buildings in the world operate at \ac{tdb} lower than this value.
Mainly because an indoor \ac{tdb} lower than \var{ta_95_perc_min} would be deemed to be unsatisfactory by a great number of occupants~\cite{iso7730}.
This lack of data in the \ac{db2} is particularly relevant for \ac{vr}.
The \ac{pmv} and \ac{pmv-ce} only differ when the value of \ac{vr}~$\geq$~\qty{0.1}{\m\per\s}.
The median value in the \ac{db2} is \var{v_median} suggesting that most of the data points were collected in buildings with low air movement.
We also could not reliably test the accuracy of the \ac{pmv} formulations for values of \ac{vr}~$\geq$~\var{v_95_perc_max}.

@@

\subsection{Comparison of PMV Accuracy in Predicting Thermal Sensation}\label{subsec:model-accuracy-comparison-in-predicting-thermal-sensation}
The \ac{pmv} was developed with the primary aim of predicting \ac{tsv}, consequently, we compared the predicted value from the two models with measured \ac{tsv}.
However, since the two models only differ when \ac{vr}~$>$~\qty{0.1}{\m\per\s};
we also present the results for a subset of the data where \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}.
In this subset of data, the differences in the outputs of the two models are more pronounced.
We set the limit to \qty{0.2}{\m\per\s} and not a higher value since, as previously shown in Figure~\ref{fig:dist_input_data}, we could only reliably test the accuracy of the models up to \ac{vr} of \var{v_95_perc_max}.

Being \ac{tsv} the value reported by participants (i.e., ground truth), in Figure~\ref{fig:bar_stacked_model_accuracy} we grouped participants' responses by \ac{tsv} and reported the simple accuracy of the different \ac{pmv} formulations over each sub-figure.
\begin{figure}[htb!]
    \centering
    \begin{subfigure}[b]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/bar_stacked_model_accuracy_0}
    \end{subfigure}
    \par\bigskip % force a bit of vertical whitespace
    \begin{subfigure}[b]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/bar_stacked_model_accuracy_0.2}
    \end{subfigure}
    \caption{The calculated \ac{pmv} value is grouped by the \ac{tsv} value reported by the participants.
    In a) and b) for the entire dataset, while in c and d for the dataset when \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}.
    In a) and c) are shown the results for the \ac{pmv} model, and in b) and d) for the \ac{pmv-ce} model.
    The number of data points in each bin of the two charts at the top (a and b) and at the bottom (c and d) is the same, and this allows us to compare the accuracy of both models.
    The overall prediction accuracy is shown above the chart, while the model accuracy of predicting participants grouped by their \ac{tsv} is shown over each bar.
    A random model would have an accuracy of \qty{14}{\percent}.}
    \label{fig:bar_stacked_model_accuracy}
\end{figure}
Both \ac{pmv} formulations underestimate the conditions at which participants are thermally dissatisfied with their environment.
The overall accuracies of the \ac{pmv} and \ac{pmv-ce} are \qty{33}{\percent} and \qty{34}{\percent}, respectively, when using the whole dataset.
The overall accuracy of both \ac{pmv} models increase to \qty{34}{\percent} and \qty{35}{\percent}, respectively when only considering entries with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}.
The \ac{pmv-ce} model outperforms the \ac{pmv} model for \ac{tsv}~=~0 and -1.
However, the \ac{pmv} model outperforms the \ac{pmv-ce} model when considering \ac{tsv}~=~1.
The number of data points in each \ac{tsv} bin is different.
\mycite{Cheung2019} grouped people based on the same heat loss or gains (same \ac{pmv} ranges), not the same \ac{tsv} as we did here, and then quantified the sensitivity of the model, i.e., how many times the \ac{pmv} model correctly predicted \ac{tsv}.
They found that the accuracy of \ac{pmv} in predicting \ac{tsv} was only \qty{34}{\percent} \mycite{Cheung2019}.

@@

In Table~\ref{tab:f1}, we report the F1 scores for both models, i) using the whole dataset, ii) only keeping entries with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}, iii) only keeping data points that have \ac{vr} measured at three heights with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}, and iv) for those with a $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
\begin{table}[htb!]
    \centering
    \input{tables/f1}
    \caption{F1 scores for the \ac{pmv} and \ac{pmv-ce} models for different subsets of data.}
    \label{tab:f1}
\end{table}
The F1-macro score, which is free from label imbalance, depicts that the accuracy of both models across all metrics is marginally better than random guessing (i.e., \qty{14}{\percent}) for all data.
For the subset of data with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} the F1-macro scores of the \ac{pmv} and \ac{pmv-ce} models only slightly improved.
We then analyzed the subset of data with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} and with \ac{vr} measured at three heights.
Both F1-macro scores decrease and the \ac{pmv-ce} (\num{0.16}) is lower than the value for the \ac{pmv}~ (\num{0.17}).
Our results show that the assumption that the \ac{pmv-ce} better takes into account the effect of elevated air speed is false.
All the F1 scores significantly increase when analyzing the dataset with $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
This highlights that the \ac{pmv} models are only accurate in predicting the thermal sensation of people when the thermal sensation is close to neutral.

@@

\subsubsection{Model Overall Bias}\label{subsubsec:model-overall-bias}
We calculated the overall bias of the model as previously done by \mycite{Humphreys2002} by subtracting the \ac{pmv} and \ac{pmv-ce} models prediction from the self-reported \ac{tsv}.
Figure~\ref{fig:hist_discrepancies} contains the summary statistics.
Overall both the \ac{pmv} and \ac{pmv-ce} models have a median bias lower than half of a thermal sensation interval.
The results highlight that the \ac{pmv-ce} has a higher bias than the \ac{pmv} both when calculating the bias for all the data and for the subset of data with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}.
We report the median and inter-quantile ranges since data are not normally distributed.
%  We cannot change into a letter value plot since in that case, it would be hard to highlight the number of responses that are within the range from -.5 to 5.
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{figures/hist_discrepancies}
    \caption{Overall bias of the \ac{pmv} and \ac{pmv-ce} models for the subset of data with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}.
    The x-axis shows the discrepancy (bias) between the predicted and self-reported \ac{tsv}.
    We have binned the data into intervals of equal width.
    We colored the bars in the range from \num{-.5} to \num{.5} to highlight the number of responses that are within this range.
    The chart also shows the median and inter-quantile ranges of the data.
    }
    \label{fig:hist_discrepancies}
\end{figure*}
When including only the data with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} the median bias of the \ac{pmv} model reduces to \var{bias_median_pmv_0.2} from \var{bias_median_pmv_0}, meaning that at higher speed the \ac{pmv} model improves.
The median bias for the \ac{pmv-ce} also decreases to \var{bias_median_pmv_ce_0.2} from \var{bias_median_pmv_ce_0} but it is higher than the one for the \ac{pmv}.

\subsubsection{Model Bias as Function of Five Input Variables}\label{subsubsec:model-bias-variable}
Here we present the model bias as a function of the five input variables \ac{tdb}, \ac{vr}, \ac{rh}, \ac{met}, and \ac{clo}.
Results are shown in Figure~\ref{fig:bias_models}.
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{figures/bias_models}
    \caption{Bias of the \ac{pmv} and \ac{pmv-ce} models plotted as a function of the input variables.
    The x ticks labels show the bin range.
    The square bracket($]$) means that the right value is included in the range.
    The red dashed lines show the acceptable range of bias.
    Each bin has a different number of data points.
    }
    \label{fig:bias_models}
\end{figure*}
In each sub-figure, we have binned the results into discrete intervals.
The x-axis tick label shows the bin range.
All the boxen-plots have the same area to aid visualization, even though the number of points in each bin is not constant.
The results depict that the use of the \ac{pmv-ce} model does not improve the bias of the \ac{pmv} results across all input variables.
On the contrary, the bias of the \ac{pmv-ce} diverges more from 0 and is significantly worse (\num{-.33} versus \num{-0.13}) than the one for the \ac{pmv} for \qty{0.3}{\m\per\s}~$<$~\ac{vr}~$\leq$~\qty{0.46}{\m\per\s}.
Both models have a median bias higher or equal to $\lvert0.5\lvert$ for \ac{tdb}~$\leq$~\qty{21}{\celsius} and \ac{clo}~$\leq$~\qty{.5}{clo}.
